---
title: "DS202 Final Project"
author: "Alison Wyatt"
date: "`r Sys.Date()`"
output: html_document
---

Team members: Alison Wyatt, Kylie Knudten

Data set: 

```{r}
ufo = read.csv("https://raw.githubusercontent.com/aawyatt/DS202-Final-Project/main/complete.csv")
```

Our question:
Is there a region of the US that has the most UFO sightings?

```{r}
library(dplyr)
library(stringr)
library(lubridate)
library(ggplot2)
```

## Step 1: Cleaning
Note: Because of the need for 10 variables, we have included one more variable from what was included in our HW 11 submission ('datetime') and joined another data set to get a new variable of state 'region'.
```{r}
# Steps for cleaning the dataset:

# 1. Filter out all countries except for country = "us". 
ufo1 = ufo %>% filter(ufo$country == "us")

# 2. Capitalize state and city values for cleanliness.
ufo1 = ufo1 %>% mutate(state = toupper(state), city = str_to_title(city))

#3. Extract just the year value from date.posted variable.
ufo1$date.posted = mdy(ufo1$date.posted)
ufo1 = ufo1 %>% mutate(date.posted = year(date.posted))

# 4. Extract just the time value from the datetime variable.
ufo1$datetime = mdy_hm(ufo1$datetime)
ufo1 = ufo1 %>% mutate(datetime = hour(datetime))

library(maps)
# states data
states = map_data("state") %>% mutate(region = str_to_title(region)) %>% rename(State = region)
# table with the states/abbreviations
abbr = read.csv("https://raw.githubusercontent.com/jasonong/List-of-US-States/master/states.csv")
# table with regions
region = read.csv("https://raw.githubusercontent.com/cphalpert/census-regions/master/us%20census%20bureau%20regions%20and%20divisions.csv")


ufo1 = ufo1 %>% rename(Abbreviation = state) %>% filter(Abbreviation != "PR")

ufo1 = right_join(ufo1, abbr, by = "Abbreviation")
# states for which we do not have map data
nomatch1 = ufo1 %>% anti_join(states, by = "State")
#unique(nomatch1$State)
# states for which we do not have ufo data
nomatch2 = states %>% anti_join(ufo1, by = "State")
#unique(nomatch2$State)
# remove data for which we don't have map/ufo data for
ufo1 = ufo1 %>% filter(!(State %in% c("Hawaii", "Alaska", "District of Columbia")))
# add state map information to ufo1
ufo1 = ufo1 %>% left_join(states, by = "State") %>% select(-order, -subregion)
# add region information to ufo1
ufo1 = ufo1 %>% left_join(region, by = "State") %>% select(-State.Code, -Division)

# get rid of unnecessary columns
ufo1 = ufo1 %>% select(-duration..hours.min., -comments)

# rename long/lat variables for clarity
ufo1 = ufo1 %>% rename(long.border = long, lat.border = lat)

ufo1
```

## Step 2: Visual summary (6 out of 8 candidate plots)

### Bar graph (state breakdown and top 5 cities)
```{r}
# Overall look at sightings by state
ufo1 %>% group_by(Region, Abbreviation) %>% count() %>% ggplot(aes(x = Abbreviation)) + geom_bar(aes(weight = n)) + theme(axis.text = element_text(size = 5)) + facet_wrap(~Region)

# Top 5 states with most sightings
ufo2 = ufo1 %>% group_by(Abbreviation) %>% summarise(n=n()) %>% arrange(desc(n))
ufo2 = ufo2[1 : 5, ]
ufo2 %>% ggplot(aes(x = Abbreviation)) + geom_bar(aes(weight = n)) + theme(axis.text = element_text(size = 5), axis.text.x=element_text(size=10),axis.text.y=element_text(size=10))

# Bottom 5 states with least sightings
ufo3 = ufo1 %>% group_by(Abbreviation) %>% summarise(n=n()) %>% arrange(n)
ufo3 = ufo3[1 : 5, ]
ufo3 %>% ggplot(aes(x = Abbreviation)) + geom_bar(aes(weight = n)) + theme(axis.text = element_text(size = 5), axis.text.x=element_text(size=10),axis.text.y=element_text(size=10))

# Top 5 cities with most sightings
ufo4 = ufo1 %>% group_by(city) %>% summarise(n=n()) %>% arrange(desc(n))
ufo4 = ufo4[1 : 5, ]
ufo4 %>% ggplot(aes(x = city)) + geom_bar(aes(weight = n)) + theme(axis.text = element_text(size = 5), axis.text.x=element_text(size=10),axis.text.y=element_text(size=10))
```
We can see that California had the most sightings by far, with North Dakota having the least. The top states with the most sightings were California, Florida, Michigan, Texas, and Washington. The 5 states with the least sightings were South Dakota, Delaware, Rhode Island, Wyoming, and North Dakota. The 5 cities with the most sightings are Houston, Seattle, Austin, Miami, and San Antonio. 3 out of the 5 cities are in Texas. The South and West regions have the most sightings. 

### Time series graph
```{r}
ufo1 %>% group_by(date.posted) %>% count() %>% ggplot(aes(x = date.posted, y = n)) + geom_line() + geom_point() +  geom_text(aes(label = date.posted))
```
This graph shows the number of sightings by year. There was the lowest recording sightings in 1998, and then a rise in sightings until it peaked in 2012 and then the number of sightings declined. 

### Map plot

```{r}
ufo1$latitude = as.numeric(ufo1$latitude)
ufo1 %>% ggplot(aes(x = long.border, y = lat.border)) + geom_path(aes(group= group)) + geom_point(aes(x=longitude,y=latitude), size=0.05, color = 'red', data = ufo1 %>% filter(longitude < 0, longitude > -130))
```

### Histogram (datetime)

```{r}
ufo1 %>% ggplot(aes(x = datetime)) + geom_histogram(bins = 10)
```
This histogram shows the number of sightings during the time of day. The values go from 0 to 23 for the 24 hours in a day. It looks like the most sightings are later in the night with the least sightings during the day when it's bright out, which makes sense. 


### Mosaic plot (regions vs shape)
```{r}
library(ggmosaic) 

ufo5 = ufo1 %>% filter(shape != "") %>% group_by(shape) %>% count() %>% arrange(desc(n))
ufo5 = ufo5[1:5, ]
ufo5$shape = factor(ufo5$shape)
ufo1%>% filter(shape %in% ufo5$shape) %>% ggplot() + geom_mosaic(aes(x=product(shape), fill=Region, weight=1))
```
This mosaic plot shows the top 5 shapes of the UFO sighting and the region it was in. We had to take the top 5 shapes that were recorded, because there were around 29 unique shapes, and that's hard to visualize. We can see that the shape 'light' had the most sightings, and the South region had the most sightings. All 4 regions had a similar distribution of the shape of the sighting.


### Boxplot(shape vs length of sighting)
```{r}
ufo1 %>% filter(shape %in% ufo5$shape) %>% ggplot(aes(x = shape, y = log(as.numeric(duration..seconds.)))) + geom_boxplot() 
```
This boxplot shows the distribution of the length of the sighting by the top 5 shapes. We had to transform the duration in seconds variable to the log of the values to be able to see the entire spread of the variable. We see that the shape 'light' has the longest duration, with the shape 'triangle' with the shortest duration. They all have very similar average lengths.

## Step 3: Analysis results and conclusions

From our graphs/visualizations, we conclude that the South and West regions of the United States has the most UFO sightings between 1998 and 2014. There are individual analyses below each graph that led us to this conclusion.