---
title: "DS202 Final Project"
author: "Alison Wyatt"
date: "`r Sys.Date()`"
output: html_document
---

Team members: Alison Wyatt, Kylie Knudten

Data set: 

```{r}
ufo = read.csv("https://raw.githubusercontent.com/aawyatt/DS202-Final-Project/main/complete.csv")
```

Our question:
Is there a region of the US that has the most UFO sightings?

```{r}
library(dplyr)
library(stringr)
library(lubridate)
```

## Step 1: Cleaning
Note: Because of the need for 10 variables, we have included two more variables from what was included in our HW 11 submission (datetime and duration..hours.min.)
```{r}
# Steps for cleaning the dataset:

# 1. Filter out all countries except for country = "us". 
ufo1 = ufo %>% filter(ufo$country == "us")

# 2. Capitalize state and city values for cleanliness.
ufo1 = ufo1 %>% mutate(state = toupper(state), city = str_to_title(city))
ufo1

#3. Extract just the year value from date.posted variable.
ufo1$date.posted = mdy(ufo1$date.posted)
ufo1 = ufo1 %>% mutate(date.posted = year(date.posted))
ufo1

# 4. Extract just the time value from the datetime variable.
ufo1$datetime = mdy_hm(ufo1$datetime)
ufo1 = ufo1 %>% mutate(datetime = hour(datetime))
ufo1
```

## Step 2: Visual summary (6 out of 8 candidate plots)

```{r}

```

## Step 3: Analysis results and conclusions

